### 2015-02-16 paper/exprs.tex
expression preservation and progress

\begin{mathpar}
  \infer[ET-REFIMM]
    {\tc{\gG}{\lv{x}{p}}{\gt} \\
     \canread{p}{\gU(x)} \\
     \valid{\gG}{\ltr}{L}{\lv{x}{p}}{\lt} \\
     \freezable{\gG}{\ltr}{\lv{x}{p}}{\lt} \\
     \textsc{record}(\gU,\lv{x}{p},(\lt,\qimm),\gU')
    }
    {\tc{\gG;\ltr;L;\gU}{\tyref{\lt}{\qimm}{\lv{x}{p}}}{\gt;\gU'}}
\end{mathpar}

@@

\begin{mathpar}
  \infer[ET-REFMUT]
  {\tc{\gG}{\lv{x}{p}}{\gt} \\
   \cw{p}{\gU(x)} \\
   \valid{\gG}{\ltr}{L}{\lv{x}{p}}{\lt} \\
   \unique{\gG}{\ltr}{\lv{x}{p}}{\lt} \\
   \textsc{record}(\gU,\lv{x}{p},(\lt,\qmut),\gU')
  }
  {\tc{\gG;\ltr;L;\gU}{\tyref{\lt}{\qmut}{\lv{x}{p}}}{\gt;\gU'}}
\end{mathpar}

@@

\subsection*{Expression Preservation}

\subsection*{Expression Progress}

>>>>

\begin{mathpar}
  \infer[ET-REFIMM]
    {\tc{\gG}{\lv{x}{p}}{\gt} \\
     \canread{p}{\gU(x)} \\
     \valid{\gG}{\ltr}{L}{\lv{x}{p}}{\lt} \\
     \freezable{\gG}{\ltr}{\lv{x}{p}}{\lt} \\
     \textsc{record}(\gU,\lv{x}{p},(\lt,\qimm),\gU')
    }
    {\tc{\gG;\ltr;L;\gU}{\tyref{\lt}{\qimm}{\lv{x}{p}}}{\tyref{\lt}{\qimm}{\gt};\gU'}}
\end{mathpar}

@@

\begin{mathpar}
  \infer[ET-REFMUT]
  {\tc{\gG}{\lv{x}{p}}{\gt} \\
   \cw{p}{\gU(x)} \\
   \valid{\gG}{\ltr}{L}{\lv{x}{p}}{\lt} \\
   \unique{\gG}{\ltr}{\lv{x}{p}}{\lt} \\
   \textsc{record}(\gU,\lv{x}{p},(\lt,\qmut),\gU')
  }
  {\tc{\gG;\ltr;L;\gU}{\tyref{\lt}{\qmut}{\lv{x}{p}}}{\tyref{\lt}{\qmut}{\gt};\gU'}}
\end{mathpar}

@@

\subsection*{Expression Preservation}
Having defined both evaluation and typing for expressions, we can finally
state the preservation lemma for expressions. Assuming all input contexts are
well-formed and given a well typed expression that evaluates to a layout,
then that layout has the same type as the expression, the resulting shadow heap
is well-formed, and the resulting heap is consistent with the resulting shadow heap.

\begin{conj}[Expression Preservation]
  If $\tc{}{H}{\gS}$, $\gG;V\vdash\gS$, $\vdash\ltr$, $\tb{\ltr}{\gU}{\gG}{L}$, 
  $\tc{V}{H}{\gU}$, $\tc{\gG;\ltr;L;\gU}{e}{\gt;\gU'}$, and $\ev{V;H}{e}{l;H'}$
  then $\tc{\gS}{l}{\gt}$, $\tc{}{H'}{\gS}$, $\tb{\ltr}{\gU'}{\gG}{L}$, and $\tc{V}{H'}{\gU'}$.
\end{conj}

\begin{proof}[Justification]
  This is trivial in the integer case.
  In the reference case, the heap does not change.
  From Expression Shadow Preservation, we get the well-formedness of the new shadow heap,
  which has been extended with one new loan. Since the intialization information in the shadow
  heap has not changed, it should still model the heap correctly.
  That the type of the layout matches the type of the expression is simple to establish.

  In the use case, we again get the well-formedness of the new shadow heap from
  Expression Shadow Preservation. The type of the layout we can get via 
  Path Type Preservation and Read Safety. The write we perform on the heap
  does not change the type of anything, so the new heap will still typecheck under the heap type.
  In the new shadow heap, the used path was deinitialized (if affine).
  In the new heap, the used path was voided (if affine).
  The new heap should be consistent with the new shadow heap.
\end{proof}

\subsubsection*{The Layout Resulting from Evaluation is Fully Initialized}
We should also note that the resulting layout is fully initialized.
This is trivial in the integer case. In the reference and the use case
we rely on typing judgments that asserted full initialization and the correspondence
of the shadow heap with the runtime heap.

\subsection*{Expression Progress}
The progress lemma for expressions is surprisingly straightforward.
Since we do not nest expressions, no induction is necessary.
Each case constructs the resulting values in a straightforward manner.

\begin{conj}[Expression Progress]
  If $\tc{}{H}{\gS}$, $\gG;V\vdash\gS$, $\vdash\ltr$, $\tb{\ltr}{\gU}{\gG}{L}$,
  $\tc{V}{H}{\gU}$, and $\tc{\gG;\ltr;L;\gU}{e}{\gt;\gU'}$
  then there exists a layout $l$ and a heap $H'$ such that $\ev{V;H}{e}{l;H'}$.
\end{conj}

\begin{proof}[Justification]
  In the integer case this is trivial.
  In both reference cases, we can rely on Path Progress to evaluate the L-value
  and use the address to construct the resulting layout, and the heap does not change.
  In the use case, we can rely on Path Progress to evaluate the L-value to an address.
  Then we can rely on Read Safety to find the layout. Since we could read sufficiently,
  we can also write sufficiently so we find the heap.

\end{proof}
